Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-07-31 Scripts/Manuscript_Draft.Rmd
updated numbers in manuscript

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 75 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear **more intensively fished** than those mainly catching non-transboundary species. Our study has direct implications **for fisheries management** of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

@@

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 243 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level **(or even as a group of species) like the case of shrimps (e.g. Litopenaeus stylirostris or Farfantepenaeus californiensis) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt], hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf].** Also, recent research has suggested strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis **while still making it relevant from a management perspective**. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only consider a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the 938 species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting transboundary species within EEZs **caught** an **annual** average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating **a yearly average of USD 75,961 million in fishing revenue during the same time period**. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 35,612 million (**converted to in 2019 USD**) in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. **Such difference could be due to the utilization of reported catch data by Teh and Sumaila instead of reconstructed data and a time period difference (Teh and Sumaila's study ends in 2006). To compare the two studies, we used the FAO global reported data identified at the species level by the *Sea Around Us* and found that in 2006, the 633 transboundary species accounted for 40.4 million tons. While this is not necessarily a direct comparison (e.g. methodology differences might still exist), it suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated due to an incomplete knowledge on the transboundary nature of marine fished species**

Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, our study shows that Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile and thus discrete for Ecuador (**spatial threshold between Ecuador and Peru < 10%**) **[@Cashion:2018cg]**. The **annual average contribution of discrete species (in our previous example would be only considering Ecuador's anchoveta catch) between 2005 and 2014** was 10.8 million tons representing USD 19,243 million in fishing revenue **over the same time period**. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary **according to our methods of determination (See methods)**. These species represented only **an annual average of** 2.8 million tons and 4,282 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (**average yearly** revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 6,279 million) in **Eastern** Asia and the USA **in Northern America** (USD 9,605 million) contribute 55% and 82% to the **annual average** revenues. In addition to China and the USA, Russia (USD 6,361 million), Peru (USD 5,211 **million**) and Japan (USD 3,368 **million**) are among the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species.

@@

**We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is not to infer stock status but to show the catch trend of transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods).** We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,447) = 27.091, *p* < 0.001). Specifically, **occasions where a transboundary species is only fished by one country are generally less likely to have a decreasing catch trend compared to when the species is shared by neighboring countries (Table S2)**. Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,447) = 55.01, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,380) = 21.86, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.052).

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that **straddles into the high seas** that contributed an average of 598.4 thousand tons to **total** catches (8.19% of total) and USD 850.2 million in **fishing** revenue (7.17% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change in the distribution of shared species is key to develop local adaptation methods to anticipate negative impacts to sustainability. **For example, understanding how climate change will modify the proportion that each neighboring EEZ has over a sharing species [@PalaciosAbrantes:2020vt], when this change is going to happen, and the economic consequence of such effects can make international treaties more resilient to climate change [@RashidSumaila:tp].**

>>>>

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 82% of the global-EEZs catches (48 million tons) and 81% of fisheries revenue (USD 77 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear **more intensively fished** than those mainly catching non-transboundary species. Our study has direct implications **for fisheries management** of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

@@

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 280 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level **(or even as a group of species) like the case of shrimps (e.g. Litopenaeus stylirostris or Farfantepenaeus californiensis) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt], hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf].** Also, recent research has suggested strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis **while still making it relevant from a management perspective**. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only consider a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the 938 species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting transboundary species within EEZs **caught** an **annual** average of 48.5 million tons, representing 82.3% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating **a yearly average of USD 77,591 million in fishing revenue during the same time period**. This is equivalent to 81.3% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 40,187 million (**converted to in 2019 USD**) in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. **Such difference could be due to the utilization of reported catch data by Teh and Sumaila instead of reconstructed data and a time period difference (Teh and Sumaila's study ends in 2006). To compare the two studies, we used the FAO global reported data identified at the species level by the *Sea Around Us* and found that in 2006, the 633 transboundary species accounted for 40.7 million tons. While this is not necessarily a direct comparison (e.g. methodology differences might still exist), it suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated due to an incomplete knowledge on the transboundary nature of marine fished species**

Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, our study shows that Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile and thus discrete for Ecuador (**spatial threshold between Ecuador and Peru < 10%**) **[@Cashion:2018cg]**. The **annual average contribution of discrete species (in our previous example would be only considering Ecuador's anchoveta catch) between 2005 and 2014** was 7.3 million tons representing USD 12,710 million in fishing revenue **over the same time period**. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary **according to our methods of determination (See methods)**. These species represented only **an annual average of** 2.8 million tons and 4,282 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (**average yearly** revenue = 4,680 $\pm$ 6,000 million USD) and Eastern Asia (3,779 $\pm$ 3,093 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,165) = 5.081, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 7,284 million) in **Eastern** Asia and the USA **in Northern America** (USD 11,604 million) contribute 55% and 82% to the **annual average** revenues. In addition to China and the USA, Russia (USD 7,379 million), Peru (USD 6,044 **million**) and Japan (USD 3,907 **million**) are among the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 41% of the global fisheries revenue from transboundary species.

@@

**We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is not to infer stock status but to show the catch trend of transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods).** We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,459) = 29.406, *p* < 0.001). Specifically, **occasions where a transboundary species is only fished by one country are generally less likely to have a decreasing catch trend compared to when the species is shared by neighboring countries (Table S2)**. Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,459) = 53.93, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,382) = 25.58, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.058).

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that **straddles into the high seas** that contributed an average of 598.4 thousand tons to **total** catches (8.18% of total) and USD 986.24 million in **fishing** revenue (7.16% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change in the distribution of shared species is key to develop local adaptation methods to anticipate negative impacts to sustainability. **For example, understanding how climate change will modify the proportion that each neighboring EEZ has over a sharing species [@PalaciosAbrantes:2020vt], when this change is going to happen, and the economic consequence of such effects can make international treaties more resilient to climate change [@RashidSumaila:tp].**
